---
title: "Model validation"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(rms)
library(dplyr)
library(tidyr)
library(caret)
library(ResourceSelection)
```

This notebook validates the prognostic model generated from EDP dataset on the DRC data.

```{r message=FALSE, cache=F}
source('./mdl-evaluation.R')
```

Fist, loading model:

```{r}
mdl_folder = "model-age-ct-symp"
load(paste0(mdl_folder, "/model.RData"))
```

Second, loading data:

```{r}
test_data_file = 'drc-data.csv' 
test_data <- read.table(test_data_file, sep=",", header=TRUE, na.strings="\\N")
  
# Renaming variables so the model can be applied on the data
test_data %>% 
  rename(
    Death = S7.FinalStatus,
    PatientAge = AD.Age,
    CT = Lab.CtNP.D1,
    AnyBleeding = S2.Bleeding,
    Diarrhoea = S2.Diarrhea,
    Breathlessness = S2.DifficultyBreathing, 
    SwallowingProblems = S2.DifficultySwallowing
    ) -> test_data

test_data %>%
 select(Death, PatientAge, CT, AnyBleeding, Diarrhoea, Breathlessness, SwallowingProblems) -> test_data

test_data %>% drop_na() -> test_data

test_data
```

Several model evaluation measures from RMS 

```{r}
dir <- mdl_folder
dat <- test_data
mdl <- model

# Important note: the function in this cell cannot be run inside a function for some reason...

dd <- datadist(dat)
options(datadist="dd")

mdl_anova <- anova(mdl)
mdl_update <- update(mdl, x=TRUE, y=TRUE)
plot(mdl_anova)
saveToTXT(mdl_anova, dir, "anova.txt")
saveToPDF(mdl_anova, dir, "anova.pdf")

mdl_val <- validate(mdl_update, B=100)
print(mdl_val)
saveToTXT(mdl_val, dir, "validation.txt")

#mdl_cal <- calibrate(mdl_update, B=100)
#plot(mdl_cal)
#saveToPDF(mdl_cal, dir, "calibration.pdf")

f <- mdl_update
plot(summary(f), log=TRUE)

pdf(paste0(dir, "/or.pdf"), useDingbats=FALSE) 
plot(summary(f), log=TRUE)  
dev.off()  
```

Generate calibration and ROC plots

```{r}
pred <- predict(model, test_data)
probs <- exp(pred) / (1 + exp(pred))
obs <- test_data$Death

data = data.frame(out = obs, prob = probs)
data$logodds = log(data$prob / (1-data$prob))
m1 <- glm(data$out ~ -1 + offset(data$logodds), family = binomial)
m1

data$m1_pred <- predict(m1, type = "response")

cal_plot(m1, data, "m1_pred", model_label ='', dir = mdl_folder, fn = "cal_plot_validation.pdf")
c1 <- cal_coeff(data, "m1_pred", mdl_folder, "cal_plot_coefficients.txt")
print(c1)
```

The plot above shows that the model is not well calibrated. The test below quantifies that:

```{r}
hl <- hoslem.test(data$out, data$prob, g=10)
print(hl)
```

The test rejects the null hypothesis of predicted and observed probabilities being the same.

The ROC plot including the AUC with CIs:

```{r}
roc_plot(obs, probs, dir = mdl_folder, fn = "roc_plot_validation.pdf")
```

```{r}
plot_roc_perf(obs, probs)
```

Finally, the confusion matrix is calculated below:

```{r}
conf_matrix(obs, probs, 0.63, mdl_folder, "conf_matrix.txt")
``` 

=============================================================================================================================

Calculate performance on reference model (age + Ct) from Colubri & Hartley 


```{r}
# Construct the reference prognostic model that only uses age and Ct in the first 48 hours

ref_model <- function(df) {
  # Calculate the linear term in the logistic function of the model
  # p(x) = 1 / (1+exp(-lp(x))
  lp = with(df, 
            0.86215 - 
              0.05343*PatientAge + 
              3*10^-5*max(PatientAge - 5.0, 0)^3 - 
              6*10^-5*max(PatientAge - 30.0, 0)^3 + 
              3*10^-5*max(PatientAge - 58.0, 0)^3 - 
              0.8694*CTnorm)
  return(lp)
}
```

```{r}
# Add normalized Ct to data, needed by reference model
test_data$CTnorm <- (test_data$CT - mean(test_data$CT, na.rm=T))/sd(test_data$CT, na.rm=T)
```

```{r}
mdl_folder0 <- "model-ref"
dir.create(mdl_folder0)

vars0 <- c('Death', 'PatientAge', 'CTnorm')
all_data0 <- test_data %>% select(all_of(vars0))
                             
data0 <- all_data0[complete.cases(all_data0),]
obs0 <- data0$Death

pred0 <- unlist(map(1:nrow(data0), function(x) ref_model(data0[x,]))) # prediction in log-odds (probabilities on logit scale)
probs0 <- 1 / (1 + exp(-pred0))                                       # prediction as probability response

# Generate the data frame with the data needed for the calibration and ROC plots
plot_data0 <- data.frame(out = data0$Death, logodds = pred0, prob = probs0)

cal_plot(ref_model, plot_data0, "prob", model_label = '', dir = mdl_folder0, fn = "cal_plot.pdf")
c0 <- cal_coeff(plot_data0, "prob", mdl_folder0, "cal_plot_coefficients.txt")
print(c0)
```

```{r}
roc_plot(data0$Death, probs0, dir = mdl_folder0, fn = "roc_plot.pdf")
```

```{r}
plot_roc_perf(obs0, probs0)
```

```{r}
conf_matrix(obs0, probs0, 0.61, mdl_folder, "conf_matrix.txt")
```