---
title: "Model update"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(Hmisc)
library(rms)
library(tidyr)
library(dplyr)
```

This notebook updates the model applying recalibration, and then adding one more predictor

```{r message=FALSE, cache=F}
source('./mdl-evaluation.R')
```

Start by loading model:

```{r}
load("model-age-ct-symp/model.RData")
mdl_folder = "model-lab"
dir.create(mdl_folder)
```

Second, loading data:

```{r}
test_data_file = 'drc-data.csv' 
all_test_data <- read.table(test_data_file, sep=",", header=TRUE, na.strings="\\N")
  
# Renaming variables so the model can be applied on the data
all_test_data %>% 
  rename(
    Death = S7.FinalStatus,
    PatientAge = AD.Age,
    CT = Lab.CtNP.D1,
    AnyBleeding = S2.Bleeding,
    Diarrhoea = S2.Diarrhea,
    Breathlessness = S2.DifficultyBreathing, 
    SwallowingProblems = S2.DifficultySwallowing
    ) -> test_data

test_data %>%
 select(Death, PatientAge, CT, AnyBleeding, Diarrhoea, Breathlessness, SwallowingProblems) -> test_data

test_data %>% drop_na() -> test_data

test_data
```

1. RECALIBRATION

The model recalibration follows the methods described in

https://darrendahly.github.io/post/homr/

1.2 Recalibration on the large 

Only the intercept is refitted:

```{r}
pred <- predict(model, test_data)
probs <- exp(pred) / (1 + exp(pred))
obs <- test_data$Death

data = data.frame(out = obs, prob = probs)
data$logodds = log(data$prob / (1-data$prob))
m2 <- glm(data$out ~ 1 + offset(data$logodds), family = binomial)
m2

data$m2_pred <- predict(m2, type = "response")

cal_plot(m2, data, "m2_pred", model_label ='', dir = mdl_folder, fn = "cal_plot_recalibration_on_the_large.pdf")
c2 <- cal_coeff(data, "m2_pred", mdl_folder, "cal_plot_recalibration_on_the_large_coefficients.txt")
print(c2)
```

1.2 Logistic recalibration

Both the intercept and slope are refitted:

```{r}
m3 <- glm(data$out ~ 1 + data$logodds, family = binomial)
m3

data$m3_pred <- predict(m3, type = "response")

cal_plot(m3, data, "m3_pred", model_label ='', dir = mdl_folder, fn = "cal_plot_recalibration_logistic.pdf")
c3 <- cal_coeff(data, "m3_pred", mdl_folder, "cal_plot_recalibration_logistic_coefficients.txt")
print(c3)
```

2. ADDING PREDICTOR TO MODEL

Applying the "recalibration with extension method described" in:

https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-016-0231-2

First, looking for the lab result that has highest correlation with outcome

```{r}
lab_data <- read.table(test_data_file, sep=",", header=TRUE, na.strings="\\N")

all_test_data %>%
 select(S7.FinalStatus, Lab.AmLyteGLU.D1.D2, Lab.AmLyteCRE.D1.D2, Lab.AmLyteALB.D1.D2, Lab.AmLyteAST.D1.D2, Lab.AmLyteAMY.D1.D2, Lab.AmLyteK.D1.D2, 
    Lab.AmLyteCRP.D1.D2, Lab.AmLyteBUN.D1.D2, Lab.AmLyteTBIL.D1.D2, Lab.AmLyteALT.D1.D2, Lab.AmLyteCK.D1.D2, Lab.AmLyteNa.D1.D2, Lab.AmLyteCa.D1.D2) -> lab_data

lab_data
```

```{r}
# ++++++++++++++++++++++++++++
# flattenCorrMatrix
# ++++++++++++++++++++++++++++
# cormat : matrix of the correlation coefficients
# pmat : matrix of the correlation p-values
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

res <- cor(lab_data)
res2 <- rcorr(as.matrix(lab_data))
res3 <- flattenCorrMatrix(res2$r, res2$P)
res3
```

```{r}
res3[res3$column == "Lab.AmLyteCK.D1.D2" & res3$row == "Lab.AmLyteALT.D1.D2",]
res3[res3$column == "Lab.AmLyteCK.D1.D2" & res3$row == "Lab.AmLyteAST.D1.D2",]
res3[res3$column == "Lab.AmLyteALT.D1.D2" & res3$row == "Lab.AmLyteAST.D1.D2",]
```

From inspecting the table below, the only lab results with significant correlation with outcome are:
Lab.AmLyteALT.D1.D2 (c=0.57, P=5.83-04), 
Lab.AmLyteAST.D1.D2 (c=0.56, P=2.8e-03), and 
Lab.AmLyteCK.D1.D2 (c=0.51, P=3.6e-03)

Let's build new models where these variables are added as a new predictor

```{r}
all_test_data %>% 
  rename(
    Death = S7.FinalStatus,
    PatientAge = AD.Age,
    CT = Lab.CtNP.D1,
    AnyBleeding = S2.Bleeding,
    Diarrhoea = S2.Diarrhea,
    Breathlessness = S2.DifficultyBreathing, 
    SwallowingProblems = S2.DifficultySwallowing,
    ALT = Lab.AmLyteALT.D1.D2,
    AST = Lab.AmLyteAST.D1.D2,
    CK = Lab.AmLyteCK.D1.D2
    ) -> test_data_labs
```

1. ALT

```{r}
test_data_labs %>%
 select(Death, PatientAge, CT, AnyBleeding, Diarrhoea, Breathlessness, SwallowingProblems, ALT) -> test_data1

test_data1 %>% drop_na() -> test_data1
test_data1
```

```{r}
pred <- predict(model, test_data1)
probs <- exp(pred) / (1 + exp(pred))
obs <- test_data1$Death

data1 = data.frame(out = obs, prob = probs)
data1$logodds = log(data1$prob / (1-data1$prob))

m1 <- glm(data1$out ~ 1 + data1$logodds + test_data1$ALT, family = binomial)
m1

data1$pred <- predict(m1, type = "response")

cal_plot(m1, data1, "pred", model_label ='', dir = mdl_folder, fn = "cal_plot_model_with_alt.pdf")
c1 <- cal_coeff(data1, "pred", mdl_folder, "cal_plot_validation_with_alt_coefficients.txt")
print(c1)

roc_plot(data1$out, data1$pred, dir = mdl_folder, fn = "roc_plot_validation_with_alt.pdf")
```

```{r}
plot_roc_perf(data1$out, data1$pred)
```

```{r}
conf_matrix(data1$out, data1$pred, 0.15, mdl_folder, "conf_matrix_model_with_alt.txt")
``` 

2. AST

```{r}
test_data_labs %>%
 select(Death, PatientAge, CT, AnyBleeding, Diarrhoea, Breathlessness, SwallowingProblems, AST) -> test_data2

test_data2 %>% drop_na() -> test_data2
test_data2
```

```{r}
pred <- predict(model, test_data2)
probs <- exp(pred) / (1 + exp(pred))
obs <- test_data2$Death

data2 = data.frame(out = obs, prob = probs)
data2$logodds = log(data2$prob / (1-data2$prob))

m2 <- glm(data2$out ~ 1 + data2$logodds + test_data2$AST, family = binomial)
m2

data2$pred <- predict(m2, type = "response")

cal_plot(m2, data2, "pred", model_label ='', dir = mdl_folder, fn = "cal_plot_model_with_ast.pdf")
c2 <- cal_coeff(data2, "pred", mdl_folder, "cal_plot_validation_with_ast_coefficients.txt")
print(c2)

roc_plot(data2$out, data2$pred, dir = mdl_folder, fn = "roc_plot_validation_with_ast.pdf")
```

```{r}
plot_roc_perf(data2$out, data2$pred)
```

```{r}
conf_matrix(data2$out, data2$pred, 0.25, mdl_folder, "conf_matrix_model_with_ast.txt")
``` 

3. CK

```{r}
test_data_labs %>%
 select(Death, PatientAge, CT, AnyBleeding, Diarrhoea, Breathlessness, SwallowingProblems, CK) -> test_data3

test_data3 %>% drop_na() -> test_data3
test_data3
```

```{r}
pred <- predict(model, test_data3)
probs <- exp(pred) / (1 + exp(pred))
obs <- test_data3$Death

data3 = data.frame(out = obs, prob = probs)
data3$logodds = log(data3$prob / (1-data3$prob))

m3 <- glm(data3$out ~ 1 + data3$logodds + test_data3$CK, family = binomial)
m3

data3$pred <- predict(m3, type = "response")

cal_plot(m3, data3, "pred", model_label ='', dir = mdl_folder, fn = "cal_plot_model_with_ck.pdf")
c3 <- cal_coeff(data3, "pred", mdl_folder, "cal_plot_validation_with_ck_coefficients.txt")
print(c3)

roc_plot(data3$out, data3$pred, dir = mdl_folder, fn = "roc_plot_validation_with_ck.pdf")
```

```{r}
plot_roc_perf(data3$out, data3$pred)
```
```{r}
conf_matrix(data3$out, data3$pred, 0.24, mdl_folder, "conf_matrix_model_with_ck.txt")
```

4. AST + CK 

```{r}
test_data_labs %>%
 select(Death, PatientAge, CT, AnyBleeding, Diarrhoea, Breathlessness, SwallowingProblems, AST, CK) -> test_data4

test_data4 %>% drop_na() -> test_data4
test_data4
```

```{r}
pred <- predict(model, test_data4)
probs <- exp(pred) / (1 + exp(pred))
obs <- test_data4$Death

data4 = data.frame(out = obs, prob = probs)
data4$logodds = log(data4$prob / (1-data4$prob))

m4 <- glm(data4$out ~ 1 + data4$logodds + test_data4$AST + test_data4$CK, family = binomial)
m4

data4$pred <- predict(m4, type = "response")

cal_plot(m3, data4, "pred", model_label ='', dir = mdl_folder, fn = "cal_plot_model_with_ast_ck.pdf")
c4 <- cal_coeff(data4, "pred", mdl_folder, "cal_plot_validation_with_ast_ck_coefficients.txt")
print(c4)

roc_plot(data4$out, data4$pred, dir = mdl_folder, fn = "roc_plot_validation_with_ast_ck.pdf")
```

```{r}
plot_roc_perf(data4$out, data4$pred)
```
```{r}
conf_matrix(data4$out, data4$pred, 0.5, mdl_folder, "conf_matrix_model_with_ast_ck.txt")
```

5. ALT + CK

```{r}
test_data_labs %>%
 select(Death, PatientAge, CT, AnyBleeding, Diarrhoea, Breathlessness, SwallowingProblems, ALT, CK) -> test_data5

test_data5 %>% drop_na() -> test_data5
test_data5
```

```{r}
pred <- predict(model, test_data5)
probs <- exp(pred) / (1 + exp(pred))
obs <- test_data5$Death

data5 = data.frame(out = obs, prob = probs)
data5$logodds = log(data5$prob / (1-data5$prob))

m4 <- glm(data5$out ~ 1 + data5$logodds + test_data5$ALT + test_data5$CK, family = binomial)
m4

data5$pred <- predict(m4, type = "response")

cal_plot(m3, data5, "pred", model_label ='', dir = mdl_folder, fn = "cal_plot_model_with_alt_ck.pdf")
c4 <- cal_coeff(data5, "pred", mdl_folder, "cal_plot_validation_with_alt_ck_coefficients.txt")
print(c4)

roc_plot(data5$out, data5$pred, dir = mdl_folder, fn = "roc_plot_validation_with_alt_ck.pdf")
```

```{r}
plot_roc_perf(data5$out, data5$pred)
```

```{r}
conf_matrix(data5$out, data5$pred, 0.5, mdl_folder, "conf_matrix_model_with_alt_ck.txt")
```

6. ALT + AST + CK 

```{r}
test_data_labs %>%
 select(Death, PatientAge, CT, AnyBleeding, Diarrhoea, Breathlessness, SwallowingProblems, ALT, AST, CK) -> test_data6

test_data6 %>% drop_na() -> test_data6
test_data6
```

```{r}
pred <- predict(model, test_data6)
probs <- exp(pred) / (1 + exp(pred))
obs <- test_data6$Death

data6 = data.frame(out = obs, prob = probs)
data6$logodds = log(data6$prob / (1-data6$prob))

m4 <- glm(data6$out ~ 1 + data6$logodds + test_data6$ALT + test_data6$AST + test_data6$CK, family = binomial)
m4

data6$pred <- predict(m4, type = "response")

cal_plot(m3, data6, "pred", model_label ='', dir = mdl_folder, fn = "cal_plot_model_with_alt_ast_ck.pdf")
c4 <- cal_coeff(data6, "pred", mdl_folder, "cal_plot_validation_with_alt_ast_ck_coefficients.txt")
print(c4)

roc_plot(data6$out, data6$pred, dir = mdl_folder, fn = "roc_plot_validation_with_alt_ast_ck.pdf")
```

```{r}
plot_roc_perf(data6$out, data6$pred)
```

```{r}
conf_matrix(data6$out, data6$pred, 0.5, mdl_folder, "conf_matrix_model_with_alt_ast_ck.txt")
```