---
title: "Data cleanup"
output:
  html_document:
    df_print: paged
---

This notebook cleans up the DRC data by removing patients with length of stay (LOS) less than 24 hours, 
inverting final status, and aggregating lab results for days 1 and 2.

```{r setup, include=FALSE}
library(dplyr)
```

```{r}
df <- read.table("drc-data-orig.csv", sep=",", header=TRUE, na.strings="")
```

```{r}
# Removing cases with LOS less than 24 hours
df <- subset(df, LOS_bin_new != "<24h")

# Inverting outcome value (1=alive, 0=dead, but opposite is needed in prog model)
df$S7.FinalStatus <- 1 - df$S7.FinalStatus

# Creating the JoinMuscle pain variable
df$S2.JoinMusclePain <- ifelse(
    ( 
        (df$S2.MusclePain == 1) |
        (df$S2.JointPain == 1)
    ),
    1,  # if condition is met, put 1
    0   # else put 0
)
```

```{r}
df[is.na(df)] <- "\\N"
write.table(df, file="drc-data-tmp.csv", sep = ",", row.names = FALSE)
```

Another step is performed manually after saving the file, which consists in aggregating lab results 
form days 1 and 2 (using the highest value if both are non-null). Also replacing out-of-range readings
by random values above or below the limit. Maybe this can be done in the future using dplyr:

https://www.marsja.se/r-add-column-to-dataframe-based-on-other-columns-conditions-dplyr/
https://stackoverflow.com/questions/54774280/plyrddply-equivalent-in-dplyr

But for the time being it was easier to just to manually since the number of rows is not too high.